|

### 2020-09-20 paper.Rmd
Some restructuring of the discussion.

## Clustering single-cell RNA-seq data sets

@@

## Classifying hand-sketches

@@

## Comparing physics experiments: PDFSense

@@

## Tuning the parameters: Pollen

@@

In this paper we have introduced the sage display, a visualisation technique for reversing
the curse of dimensionality in low dimensional projections of high dimensional data, and have shown its utility in applications in the natural sciences and machine learning. Our case studies highlight that the radial transformation can reduce the effects of data piling, and in combination with the grand tour our new sage display provides interpretable visualisation, while preserving global structure in high dimensional space.

The sage display is easily incorporated into exploratory high dimensional data analysis
tasks: classification, clustering, and search. For classification tasks,
as shown in the sketches example, the sage display decreases the number of 
overlapping points between classes and provides better visual separation between
classes compared to the regular tour.  For clustering tasks, as shown in the 
single cell RNA-seq example, the sage display uncovered clusters that were 
originally obscured by data piling. Furthermore, unlike some
non-linear dimension reduction (NLDR) techniques like t-SNE, the sage display 
still gives the viewer an accurate assessment of the size of a cluster, 
and their relative orientation; we learned that yellow and blue clusters have
less separation in high dimensional space than the clustering algorithm suggests.
For search tasks, as shown in the pollen and pdfsense examples, the sage display
allows to find hidden signal that is concealed by the density of points around
the center of the projection. In the pdfsense example, by zooming in on the 
points in the blue cluster we saw how they were present in all directions of the
parameter space, which was difficult to see with the regular tour.

The applications also highlight that tuning is often required,
especially if the projection is dense in the center of the 
distribution. Expanding around the center aggressively can introduce distortions
in the display, a well known problem for other NLDR methods as well.
However, unlike those methods, the parameters in the sage display are 
interpretable; we know the effect of altering $\gamma$ and $R$ on the resulting 
visualisation. Recall that the radial transformation is approximately linear near
the center, thus preserving structures such as those seen in the pdfsense and
pollen applications. On the other hand, the tuning example showed how points away
from the projected center get moved to the edge of the hypersphere as $\gamma$
is increased or $R$ is decreased. 
<!--
Our parameter tuning example highlights this: the sage display is 
linear up to a  given effective dimension; the default sage display, preserves 
the correlation between the variables which is broken once points get moved
to the edge of the hypersphere as $\gamma$ and $R$ are changed.
-->
Given
that the display is quick to compute, being able to modify these parameters in 
real time with an interactive interface would be beneficial.

Currently, the 
sage display is naive to information in high dimensional space that is not 
captured by linear projections. Like t-SNE, it would be interesting to explore
transformations of projections where points in the high dimensional space are 
locally dense instead of expanding from the center of the distribution.

The recently introduced slice tour [@sliceTour] is essentially addressing the same issue, i.e. the overplotting of points from the large orthogonal space, but follows a different philosophy. Instead of mapping orthogonal volume to projected volume, we cut through the orthogonal space and only select a small portion of the volume for plotting. This is done by defining a slice via a cutoff on the orthogonal distance of points. The approach would be preferred in the case of large number of points, or for concave structure in the distribution, while the sage display works better in the case of limited sample size.

>>>>

## Clustering single-cell RNA-seq data sets {#sec:appl1}

@@

## Classifying hand-sketches {#sec:appl2}

@@

## Comparing physics experiments: PDFSense {#sec:appl3}

@@

## Tuning the parameters: Pollen {#sec:appl4}

@@

In this paper we have introduced the sage display, a visualisation technique for reversing
the curse of dimensionality in low dimensional projections of high dimensional data, and have shown its utility in applications in the natural sciences and machine learning. 
The new method combines animated linear dimension reduction obtained with a grand tour and a non-linear transformation of the projected data points to improve intelligibility of each view.
Our case studies show how it reduces the effects of data piling and aids different tasks in exploratory high dimensional data analysis.

The approach  provides interpretable visualisation that captures high dimensional information and preserves global structure, and it is complementary to non-linear dimension reduction techniques. For example, when visualising clusters, the sage display enables an assessment of cluster sizes, and accurately captures relative position and orientation. However, the sage display is currently 
naive to information in high dimensional space that is not 
captured by linear projections. Like t-SNE, it would be interesting to explore
transformations of projections where points in the high dimensional space are 
locally dense instead of expanding from the center of the distribution.

XXX option to reference upcoming work in liminal paper when mentioning complementarity

While the recently introduced slice tour [@sliceTour] is essentially addressing the same issue as the sage display (i.e. overplotting of points from the orthogonal space), it follows a different philosophy: we only select a small portion of the volume for plotting, based on the orthogonal distance of points from the projection plane. This approach is preferred for large sample sizes, or for concave structure, while the sage display works better in the case of limited sample size.

The sage display is easily incorporated into exploratory high dimensional data analysis, and applications shown in Section \ref{sec:application} provide examples of the following tasks:

- classification tasks: the sage display decreases the number of overlapping points between classes and provides better visual separation between classes compared to the regular tour, as shown in the sketches example (Section \ref{sec:appl2})
- clustering tasks: the sage display uncovered clusters that were originally obscured by data piling, while still giving the viewer an accurate assessment of the size of a cluster, and their relative orientation, as shown in the single cell RNA-seq example (Section \ref{sec:appl1})
- search tasks: the sage display allows to find hidden signal that is concealed by the density of points around the center of the projection, as shown in the pollen example (Section \ref{sec:appl4})

XXX is pdfsense example of search task? maybe understanding structures across multiple dimensions instead?
In the pdfsense example, by zooming in on the 
points in the blue cluster we saw how they were present in all directions of the
parameter space, which was difficult to see with the regular tour.

The applications also show that tuning is often required,
especially if the projection is dense in the center of the 
distribution. Expanding around the center aggressively can introduce distortions
in the display, a well known problem from non-linear dimension reduction techniques like t-SNE.
However, unlike those methods, the parameters in the sage display are 
interpretable; we know the effect of altering $\gamma$ and $R$ on the resulting 
visualisation. Recall that the radial transformation is approximately linear near
the center, thus preserving structures such as those seen in the pdfsense and
pollen applications. On the other hand, the tuning example showed how points away
from the projected center get moved to the edge of the hypersphere as $\gamma$
is increased or $R$ is decreased.

Given
that the display is quick to compute, being able to modify these parameters in 
real time with an interactive interface would be beneficial.
